Implications of uniformly distributed, empirically informed priors for phylogeographical model selection: A reply to Hickerson et al.

### 2014-07-30 manuscripts/reply/reply-si-main.tex
Finishing draft of reply ms overhaul.

%CWL: This seems to specific, not sure about including
\section{Validation analyses}
% \highLight{Not sure where to put this section}.
% \highLight{This justifies presenting the unadjusted results in this paper.}
% \highLight{Is it worth including just for this purpose?}
Following \citet{Oaks2012}, we characterize the model-choice behavior of the
model-averaging approach of \citet{Hickerson2013} under the ideal conditions
where the prior is correct (i.e., the data are generated from parameters drawn
from the same prior distributions used in the analysis).
We used the same prior models as above ($M_1$--$M_5$;
Table~\ref{tabModelChoiceEmpirical}), and simulated 50,000 datasets under this
prior (10,000 from each model).
We used a simulated data structure of eight population pairs, with a single
1000 base-pair locus sampled from 10 individuals from each population.
We then analyzed each of these replicate datasets using the same prior with 2.5
million samples (500,000 from each of the five prior models), retaining 1000
posterior samples.
Our results are very similar to \citet{Oaks2012}, but we note that they
are not directly comparable as our simulations contained eight population
pairs rather than 10 (Figure~\ref{figValidationMCBehavior}).
We find that the approach of \citet{Hickerson2013} estimates the posterior
probability of divergence models reasonably well when all assumptions of the
method are met (i.e., the prior is correct) and the unadjusted posterior
estimates are used.
Similar to \cite{Oaks2012}, we find that the regression-adjusted estimates of
the model probabilities are biased.

>>>>

\section{Validation analyses}
Following \citet{Oaks2012}, we characterize the model-choice behavior of the
model-averaging approach of \citet{Hickerson2013} under the ideal conditions
where the prior is correct (i.e., the data are generated from parameters drawn
from the same prior distributions used in the analysis).
We used the same prior models as above ($M_1$--$M_5$;
Table~\ref{tabModelChoiceEmpirical}), and simulated 50,000 datasets under this
prior (10,000 from each model).
We used a simulated data structure of eight population pairs, with a single
1000 base-pair locus sampled from 10 individuals from each population.
We then analyzed each of these replicate datasets using the same prior with 2.5
million samples (500,000 from each of the five prior models), retaining 1000
posterior samples.
Our results are very similar to \citet{Oaks2012}, but we note that they
are not directly comparable as our simulations contained eight population
pairs rather than 10 (Figure~\ref{figValidationMCBehavior}).
We find that the approach of \citet{Hickerson2013} estimates the posterior
probability of divergence models reasonably well when all assumptions of the
method are met (i.e., the prior is correct) and the unadjusted posterior
estimates are used.
Similar to \cite{Oaks2012}, we find that the regression-adjusted estimates of
the model probabilities are biased.

\section{A difficult inference problem}
In the main text, we discuss how the prior assumption of uniformly distributed
divergence times in \msb leads to posteriors that are difficult to interpret.
However, it is also important to consider the difficult inference problem with
which \msb is faced.
When applying \msb to the dataset of \citet{Oaks2012} with 22 taxon
pairs, there are 581--602 free parameters that model highly stochastic
coalescent and mutational processes.
Under this rich stochastic model, the method is estimating the
probability of 1002 divergence models \citep[i.e., the number of integer
partitions of $Y=22$;][]{Oaks2012}.
Furthermore, all the information in the sequence alignment of each taxon pair
is distilled into four summary statistics.
This gives us a total of 88 summary statistics (four from each of the 22 taxon
pairs) that contain minimal information about many of the $\approx 600$
parameters in the model.
More summary statistics can be used in \msb, but most are highly correlated
with the four default statistics, and thus contribute little additional
information about the parameters from the sequence data.
The large number of parameters and divergence models relative to the amount of
information in the data is undoubtedly another reason the method lacks
robustness to prior conditions.
